A Constraint Solving Approach to Parikh Images of Regular Languages

### 2023-06-05 introduction.tex
Introduction

The key ideas of \Calculus{} is a combination of problem-aware case splitting,
lazy enforcement of automata connectivity, and lazy computation of products.

We implement \Calculus{} as a theory for the \Princess{} automated theorem
prover in the tool \Catra. \Catra{} also supports the approximate method of
\cite{approximate-parikh}, its fall-back variant adapted
from~\cite{generate-parikh-image}, and an adapter for the \Nuxmv{} model
checker~\cite{nuxmv}. Using \Catra, we compare \Calculus{} to the other two
back-ends on \NrBenchmarks{} Parikh automata intersection problems generated by
\OstrichPlus{} when solving the PyEx string constraint benchmark suite involving
string length constraints \cite{pyex}.

@@

In this section we will introduce an intuition for how a string constraint problem in an automata-based solver like \OstrichPlus{} is translated to a Parikh automata intersection problem and solved efficiently using \Calculus{}. First, by a Parikh automaton we mean a standard Non-deterministic finite automaton (NFA) with a set of Parikh registers that are incremented (or decremented) at each transition. In the automata of \cref{fig:examples}, we will use the notation $a / \begin{bmatrix} x_+ \end{bmatrix}$ to mean that a transition reads an input character $a$ and increments register $x$ by one. We will omit any zero-valued increments and assume that all registers are scoped to their automaton. The increments are in reality a vector (hence the notation), but as it is mostly sparse here, we instead use the symbolic notation $\begin{bmatrix} x_{1+} \end{bmatrix}$ rather than the more cumbersome $\begin{bmatrix} 0 \\ 1 \\ 0 \\  \vdots \\ 0 \end{bmatrix}$.

@@

\begin{example}\label{ex:string-constraints}
    Consider the following set of string constraints:
\begin{constraints}
    \item\label{const:x-in-c-dd} $x \in \Regex{c\RegexOr{}dd}$
    \item\label{const:y-in-b} $y \in \Language(B)$, the language accepted by automaton~$B$ of \cref{fig:aut_b}.
    \item\label{const:x-substring} $x = \Substring(y, l, r)$
    \item\label{const:more-a-than-b} $\CountOf{\Char{a}}(y) > \CountOf{\Char{b}}(y)$, that is $y$ contains more letters \Char{a} than letters \Char{b}.
\end{constraints}
\end{example}

\cref{const:x-in-c-dd,const:y-in-b,const:x-substring} can be translated to the product of Parikh automata $B \times A'$, where $A'$ (\cref{fig:aut_a}) is the Parikh automata obtained by applying the encoding of $\Substring$ described in \cite{ostrich-plus}. Here follows an intuition for our method on $B \times A'$ under the global constraint \cref{const:more-a-than-b}. Note that the cost registers of \cref{fig:aut_b} are simply counting letters.

Our approach to the computation of the Parikh image (and therefore the solution to the intersection problem posed by \cref{ex:string-constraints}) consists of two applications of laziness; late computation of products of automata, and lazy enforcement of connectivity constraints. Both are in contrast to the eager approach to finding the Parikh image described in~\cite{generate-parikh-image}.

@@

We start by adding linear constraints requiring transition variables to represent a flow through either automata by requiring that the number of incoming transitions is equal to the number of outgoing ones, and use the automated theorem prover to annotate each transition $\Transition$ with a symbolic representation of its associated integer variable:
\begin{figure}[h]
    \centering 
  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \includegraphics[scale=\autscale]{a_annotated}
    \caption{Note: unused self-loops at $I_{A'}$ and $F_{A'}$ have been pruned for readability.}\label{fig:aut_a_annotated}
      %\Description[SHORT]{LONG}
  \end{subfigure}
  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \includegraphics[scale=\autscale]{b_annotated}
    \caption{}\label{fig:aut_b_annotated}
      %\Description[SHORT]{LONG}
  \end{subfigure}
  \caption{The starting automata, without their counter increments and with their transition variables in symbolic form.}\label{fig:propagated}
\end{figure}

Note that several transitions are already marked as unuseable (have a transition value of $0$). This includes any transition for the letter $\Char{d}$, as all counter increments for that variable in $B$ is zero and hence it must also be zero in the product of both automata. By removing these transitions to obtain new automata and computing their product, we obtain the following:

@@

By putting off computing the product $B \times A'$ until after performing linear reasoning on the number of times each transition must be taken, we have constructed a much smaller product than we would have originally. Moreover, since this automaton is a stick, all runs through it can be described by the flow-preservation equations described above without the need for costly connectivity-enforcing constraints as in~\cite{generate-parikh-image}. This illustrates the advantage of our second laziness: deferred enforcement of connectivity constraints.

>>>>

The key ideas of \Calculus{} is a combination of problem-aware case splitting,
lazy enforcement of automata connectivity, and lazy computation of products. In
essence, we use constraint-based methods to efficiently enforce the constraints
associated with computing Parikh images of regular languages.

We implement \Calculus{} as a plug-in theory for the \Princess{} automated
theorem prover. \Fudge{Others} have integrated it into the similarly
\Princess{}-backed string constraint solver \OstrichPlus{}, where it is used to
handle constraints that use cost-enriched automata, as described in
\cite{ostrich-plus}.

We additionally wrap the Parikh image solver as a stand-alone tool, \Catra.
\Catra{} also supports a variant of the approximate method of
\cite{approximate-parikh}, its fall-back variant adapted
from~\cite{generate-parikh-image}, and an adapter for the \Nuxmv{} model
checker~\cite{nuxmv}. Using \Catra, we compare \Calculus{} to the other two
back-ends on \NrBenchmarks{} distinct Parikh automata intersection problems
generated by \OstrichPlus{} when solving the PyEx string constraint benchmark
suite involving string length constraints \cite{pyex}.

@@

In this section we will introduce an intuition for how a string constraint
problem in an automata-based solver like \OstrichPlus{} is translated to a
Parikh automata intersection problem and solved efficiently using \Calculus{}.
First, by a Parikh automaton we mean a standard non-deterministic finite
automaton (NFA) with a set of Parikh registers that are incremented (or
decremented) at each transition. In the automata of \cref{fig:examples}, we will
use the notation $a / \begin{bmatrix} x_+ \end{bmatrix}$ to mean that a
transition reads an input character $a$ and increments register $x$ by one. We
will omit any zero-valued increments and assume that all registers are scoped to
their automaton. The increments are usually represented as a vector (hence the
notation), but as it is mostly sparse here, we instead use the symbolic notation
$\begin{bmatrix} x_{1+} \end{bmatrix}$ rather than the more cumbersome $\left[
0, 1,  0 , \ldots, 0 \right]$.

@@

We use standard PCRE regular expressions here and throughout the paper, writing
them $\Regex{\texttt{in brackets}}$ to make that apparent. As is common,
$\RegexOr{}$ is alternation, $\mathtt{*}$ the Kleene star, and $\mathtt{.}$,
matches any single character.

\begin{example}\label{ex:string-constraints} Consider the following set of
    string constraints:
\begin{constraints}
    \item\label{const:x-in-c-dd} $x \in \Language(\Regex{c\RegexOr{}dd})$.
    \item\label{const:y-in-b} $y \in \Language(B)$, the language accepted by
    automaton~$B$ of \cref{fig:aut_b}.
    \item\label{const:x-substring} $x = \Substring(y, i, n)$, that is $x$ is an
    $n$-length substring of $y$ starting at offset $i$.
    \item\label{const:more-a-than-b} $i = n$, that is the substring is as long as its starting offset.
\end{constraints}
\end{example}

\cref{const:x-in-c-dd,const:y-in-b,const:x-substring} can be translated to the
product of Parikh automata $B \times A'$, where $A'$ (\cref{fig:aut_a}) is the
Parikh automata obtained by applying the encoding of $\Substring$ described in
\cite{ostrich-plus}. That description can then be used to obtain values for
$\CountOf{\Char{a}}(y)$ and $\CountOf{\Char{b}}(y)$. Here follows an intuition
for our method on $B \times A'$ under the global constraint
\cref{const:more-a-than-b}. Note that the cost registers of \cref{fig:aut_b} are
simply counting letters, while the extra registers $l, r$ in $A'$ also keeps
track of the matched substring. However, these registers can be mostly
disregarded here, since we have no constraints on their values.

Our approach to the computation of the Parikh image (and therefore the solution
to the intersection problem posed by \cref{ex:string-constraints}) consists of
two applications of laziness; late computation of products of automata, and lazy
enforcement of connectivity constraints. Both are in contrast to the eager
approach to finding the Parikh image described in~\cite{generate-parikh-image},
that would have started by computing the product $A' \times B$, translated it to
a Presburger formulae with $\CountOf{\Char{a}}$ etc as free variables, and then
plugged in the constraint \cref{const:more-a-than-b}.

@@

\begin{figure}[h]
    \centering 
  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \includegraphics[scale=\autscale]{a_annotated}
    \caption{ $A'$ with its associated transition variables in symbolic form.
    Note: unused ($0$) self-loops at $I_{A'}$ and $F_{A'}$ have been pruned for
    readability.}\label{fig:aut_a_annotated}
      %\Description[SHORT]{LONG}
  \end{subfigure}
  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \includegraphics[scale=\autscale]{b_annotated}
    \caption{$B$ with its associated transition variables in symbolic forms.
    Note that state $B$ is now completely dead.}\label{fig:aut_b_annotated}
      %\Description[SHORT]{LONG}
  \end{subfigure}
  \caption{The starting automata, without their counter increments and with
  their transition variables in symbolic form.}\label{fig:propagated}
\end{figure}

We proceed by adding linear constraints requiring transition variables to
represent a flow through either automata by requiring that the number of
incoming transitions is equal to the number of outgoing ones. E.g. state $B$
would have the sum $\cancel{x_{\Tuple{B, \Char{c}, B}}} + x_{\Tuple{I_B,
\Char{b}, B}} = \cancel{x_{\Tuple{B, \Char{c}, B}}}  + x_{\Tuple{B, \Char{c},
F_B}} + x_{\Tuple{B, \Char{b}, I_B}}$ where we let $x_{\Tuple{q, l, c'}}$ refer
to the integer variable associated with a transition from state $q$ to state
$q'$ with label $l$. Note that the self-loop cancels itself out. This means that
for automata where a loop can become unreachable, these constraints do not fully
capture the constraints posed by the automaton, but requires additional
reachability constraints.

Performing symbolic reasoning on these equations and plugging the corresponding
symbolic representation of the constraints on each corresponding integer
variable back as an annotation on its transition, we get obtain the two automata
in \cref{fig:propagated}.

Several transitions are already marked as unuseable (have a transition variable
value of $0$). This includes any transition for the letter $\Char{d}$, as
$\CountOf{\Char{d}}(B) = \sum_{\Transition} 0 = 0$, since the register for
$\Char{d}$ is zero for all transitions in $B$. By removing such transitions to
obtain new automata and computing their product, we obtain the product automaton
of \cref{fig:product}.

@@

By putting off computing the product $B \times A'$ until after performing linear
reasoning on the number of times each transition must be taken and using that to
prune the terms, we have computed a smaller product than we would have with a
na\"ive approach. Moreover, since this automaton contains only one loop along
one main path, all runs through it can be described by the flow-preservation
equations described above without the need for costly connectivity-enforcing
constraints as in~\cite{generate-parikh-image}. This means that we can now fully
represent all possible solutions to \cref{ex:string-constraints} by the linear
equation $\CountOf{\Char{a}} = 2 \land \CountOf{\Char{c}} > 0 \land
\CountOf{\Char{b}} = 0 \land \CountOf{\Char{d}} = 0$, derived from applying the
same flow encoding as before.

### 2023-06-05 main.tex
Introduction

\usepackage{amsmath,empheq,fancybox}
\usepackage{paralist}
\usepackage{url}
\usepackage{color}
\usepackage{textcomp,listings}
\usepackage{array}
\usepackage{mymacros}
\usepackage{microtype}
\usepackage{listings}
\usepackage{csquotes}
\usepackage{proof}
\usepackage[capitalise]{cleveref}
\usepackage{algorithm2e}      
\usepackage{multirow}
\usepackage{mathpartir}
\usepackage{amsthm}
\usepackage{numprint}
\usepackage{ebproof}

@@

\section{A demonstration of our approach}\label{sec:motivation}

To illustrate the key features of \Calculus, we will informally present our
reasoning. We begin by considering the following set of string and other
constraints, which we want to solve for all possible values of $\MonoidElement$,
a value in some commutative monoid and the string $s$, where
$\varphi(\MonoidElement)$ is some set of constraints on $\MonoidElement$:
\[
   s \in \Language(\mathtt{/}\SndRegex{}\mathtt{/})  \land
   s \in \Language(\mathtt{/}\FstRegex{}\mathtt{/})  \land
   \MonoidElement = f(s)  \land
   \varphi(\MonoidElement)
\]

We will represent the strings using the nondeterministic finite automata
 $\SomethingCSomething{}$ (\cref{fig:something-c-something}) and $\AcaOrBc{}$
 (\cref{fig:aca-or-bc}) respectively.

 Throughout these examples we will present our three key ideas:
 \emph{automata-aware splitting}, \emph{lazy enforcement of automata
 connectivity}, and \emph{incremental materialisation of products}.

\subsection{Counting String Lengths}

To keep things simple, in this first example we let $f$ be the function that
gives the length of a given string; $f\left(\text{curious and curiouser
still}\right) = 27$, and $f\left(\text{homo}\right) +
f\left(\text{morphism}\right) = f\left(\text{homomorphism}\right) = 12$, and we
let $\varphi(\MonoidElement) = \exists k \HoldsThat m = 2k + 1 \land k \geq 0$,
i.e. require that the length of the string is odd.

\begin{figure}[p]
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[scale=\autscale]{aca_or_bc}
    \caption{An automaton recognising the regular expression
      $\FstRegex{}$.}\label{fig:aca-or-bc}
      \Description[An NFA recognising $\FstRegex{}$]{An NFA recognising $\FstRegex{}$. It has 4 states.}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.49\textwidth} 
    \includegraphics[scale=\autscale]{something_c_something}
    \caption{An automaton recognising the regular expression
    $\SndRegex{}$.}\label{fig:something-c-something}
    \Description[An NFA recognising $\SndRegex{}$]{An NFA recognising $\FstRegex{}$. It has 2 states.}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[scale=\autscale]{aca_or_bc_simplified}
    \caption{$\AcaOrBc{}$ with its symbolic transition counters after
    simplification of their equations.}\label{fig:example-simplify-1}
    \Description[The automaton $\AcaOrBc{}$ labelled with variables counting the number of times each transition is taken, after having simplified the equations.]{The starting state shows an initial flow of 1, with two outgoing transitions to $A$ ($1-l_c'$ on letter a, and $l_b' + l_c'$ on b respectively). $A$ has a flow of $l_c$, and its outgoing transition into the final state, $F$ is the same as the incoming one from $S$, the initial state. The $B$ transition has an outgoing transition labeled $c/ l_c'$ to the final state $F$, and one backwards transition $b / l_b'$ to the starting state.}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[scale=\autscale]{something_c_something_length_simplified}
    \caption{$\SomethingCSomething{}$ after using linear reasoning over lengths to replace a variable.}\label{fig:example-length-reduced}
    \Description[The automaton $\SomethingCSomething{}$ after reasoning over lengths.]{The same automaton as before for $\SomethingCSomething{}$, but with the transition from start to final annotated with a constant number 1, and the leftmost self-loop transition replaced with the equation $1 + l_c + 2l_b' -
    r_\Sigma'$.}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.49\textwidth}
      \includegraphics[scale=\autscale]{aca_or_bc_length_split_1}
      \caption{$\AcaOrBc{}$ at the split where $\FromLabelTo{S}{a}{A}$ is not
      used.}\label{fig:aca-or-bc-length-split-1}
      \Description[The automaton \AcaOrBc{} where we have removed the transition
      from $S$ to $A$.]{ The state $A$ hangs free without incoming transitions,
      and only the lower part of the automaton remains.}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.49\textwidth}
      \includegraphics[scale=\autscale]{aca_or_bc_length_split_2}
      \caption{$\AcaOrBc{}$ at the split where $\FromLabelTo{S}{a}{A}$ \emph{is} used.}\label{fig:aca-or-bc-length-split-2}
      \Description[$\AcaOrBc{}$ with the transition from $B$ to $F$ removed.]{$\AcaOrBc{}$ when we keep the transition from $S$ to $A$. This removes the transition between $B$ and $F$, but otherwise maintains the automaton.}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.49\textwidth}
      \includegraphics[scale=\autscale]{length_split_product}
      \caption{The product of $\AcaOrBc$ and $\SomethingCSomething$ after
      case splitting and removal of dead states.}\label{fig:length-split-product}
      \Description[$\AcaOrBc{}$ with the transition from $B$ to $F$ removed.]{$\AcaOrBc{}$ with the transition from $B$ to $F$ removed. There now remains only one path to an accepting state.}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.49\textwidth}
      \includegraphics[scale=\autscale]{aca_or_bc_simplified_3}
      \caption{$\AcaOrBc{}$ after further constraint propagation.}\label{fig:example-simplify-3}
      \Description[$\AcaOrBc$ without the lower section.]{$\AcaOrBc$ with only its upper section left and the $a$ transitions now set to static 1.}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.6\textwidth}
        \includegraphics[scale=\autscale]{aca_or_bc_times_something_c_something}
        \caption{$\AcaOrBc{} \times \SomethingCSomething{}$ under the constraint
        $\TransitionCount{a} > \TransitionCount{b}$.}\label{fig:final-example} \Description[The
        final product, still a boring stick automaton with only one loop for an
        arbitrary number of c's.]{A product of the intermittent automata, which
        accepts \lstinline{ac.*ca}}
  \end{subfigure}
  \caption{The collection of automata we use as running examples in this paper, with annotations illustrating intermittent states of our calculus.}    

The na\"ive approach would be to first compute the product of the automata, then
use the approach from \cite{generate-parikh-image} to find its image under $f$,
and then solve the constraints in an automated theorem prover. However, we can
take advantage of both the constraints and definitions of $f$ to prune the
automata before computing the product and help reduce the blow-up.

Similar to \cite{generate-parikh-image} we associate each transition
$\Transition$ with an integer variable $\TransitionVar_\Transition \geq 0$
representing how many times $\Transition$ is used that we leave implicitly
existentially quantified. These variables are shown as annotations under the
transition labels in \cref{fig:aca-or-bc,fig:something-c-something}. We then
apply graph-like flow reasoning by adding linear equations (e.g. $1 + {l_b}' =
l_a + l_b$, or $l_a + l_c' = 1$) that equate the incoming transitions into a
state to the ones going out. For the initial and accepting states we add an
incoming and outgoing flow of 1, respectively. We will use these linear
equations to perform reasoning to reduce the size of the product before we
compute it. Note that we have already performed some linear reasoning to
eliminate some of the variables in the graphs above, for example to introduce
the constant number $1$ in \cref{fig:something-c-something}, which comes from $1
+ r_\Sigma = x + r_\Sigma$ solved for $x$.

We sum the individual transition variables to get the length $\MonoidElement$ we
are actually interested in. For example the image of $f$ on
$\SomethingCSomething$ is given by $\exists_{r_\Sigma, r_\Sigma'} \: 2k + 1 =
r_\Sigma + 1 + r_\Sigma' \land k \geq 0$. For $\SomethingCSomething{}$, this is
all we need to produce the image since all transitions are always accessible
from any path from the initial state to the accepting state. The difficult
connectivity constraint is not necessary to enforce here, since any path through
$\SomethingCSomething{}$ is connected, and therefore the flow equations
described above are sufficient to capture its image.

The story gets more complicated with $\AcaOrBc{}$, where we have a choice of two
paths, up or down, and where that choice affects the possibility of taking
either of the two loops. In that case, we would have the image $2l_a + l_c + l_c' + l_b + l_b'$.
If we apply standard linear reasoning on the equations (e.g. $1 + {l_b}' = l_a +
l_b$, or $l_a + l_c' = 1$), we can reduce the number of bound
($\exists$-quantified) variables somewhat, obtaining the relations in
\cref{fig:example-simplify-1}, and the corresponding over-approximation on
the $f$-image $2 + l_c + 2l_b'$. It is an over-approximation because it only
considers the number of times each transition is taken and the existence of a
path from an initial to an accepting state, but omits the connectivity of that
path in the presence of loops. For example, no valid path would have both $l_c >
0$ and $1 - l_c' = 0$, but no such restriction exists in the formula above.

Still, this approximate view allows us to perform further reasoning. Since we
know that the length of a string accepted by the intersection of the two
automata's languages must be the same, we also know that their images must be
the same under $f$, e.g. that $r_\Sigma + 1 + r_\Sigma' = 2 + l_c + 2l_b' \iff
r_\Sigma + r_\Sigma' = 1 + l_c + 2l_b'$. This in turn allows us to replace one
additional variable, e.g. $r_\Sigma = 1 + l_c + 2l_b' - r_\Sigma'$ (see
\cref{fig:example-length-reduced}). Note that the two automata are now in
contact, allowing us to propagate reasoning on one to the other.

Since we cannot take a transition a negative number of times, we have an
implicit constraint on any transition variable that it must be at least $0$.
This means that we have $1 + l_c + 2l_b' - r_\Sigma'$, implying an upper bound
on $r_\Sigma'$: $r_\Sigma' \leq 1 + l_c + 2l_b'$.

This is as far as we can get with this reasoning. We must now choose: either we
compute (materialise) the product and continue our reasoning, or we perform a
case split. Let us try a case split in order to put off computing a potentially
large product.

We apply the principle of splitting and select an early transition in the
largest automaton with loops, in our case $\FromLabelTo{S}{a}{A}$ of $\AcaOrBc$.
This creates two cases: $1-l_c' = 0$ and $1-l_c' > 0$. We start with the first
case, and after propagating the now known values and removing any transition
that becomes zero, we get the automaton of \cref{fig:aca-or-bc-length-split-1}.

We immediately notice that the self-loop of state $A$ is now disconnected from
the initial state and must be removed. This condition can be detected
efficiently using standard forwards/backwards reachability from the initial
and accepting states respectively. In other words, $2k+1 = 0$ for this case
and is implied by the splitting constraint. However, no choice of integer $k$
will satisfy this equation. Therefore, we close this branch and proceed with
the other one -- where $1-l_c' > 0$. Note that we do so without ever computing
a potentially expensive product of automata.

Plugging in the assumption $1-l_c' > 0$ (and its consequence that $l_c' = 0$),
we also get a reduced variant of $\AcaOrBc$, as seen in
\cref{fig:aca-or-bc-length-split-2}. Note that there is now no way to select any
transition variable to disconnect a loop from a path between $S$ and $F$,
and we can therefore stop worrying about enforcing connectivity at all.

Before continuing, we need to tie the inequalities for the transitions of the
automata of the product to the new transition counting variables we introduced
in the product automaton to ensure that they are consistent. In our case, we
obtain the following equations from the principle that a transition variable in
either automaton must be equal to the sum of variables in the resulting edges of
the product, which gives the following two interesting inequalities from
$\AcaOrBc{}$'s transition $\FromLabelTo{A}{c}{A}$: $2k + 1 = r_1 + r_2 + 1$

Note how this conserves the requirement that the number of loop iterations be
odd even when the same original loop appears three times in the product, twice
as a self-loop and once as a regular state transition. Using the same principle
as before for computing the length by adding up the transition counters, we
obtain the quantified formula $1 + 2k + 1 + 1 + r_b + r_b= 3 + 2k + 2r_b$ for
the length. Eliminating the final quantifier gives us the image of $f(\AcaOrBc
\times \SomethingCSomething)$ modulo odd lengths: $3 + 2k$, for some integer
$k$.

\subsection{Counting Letters}\label{sec:introduction:parikh}

Now assume that rather than the length (e.g. the count of the characters) we
want to find the whole Parikh image of the intersection of $\AcaOrBc{}$ and
$\SomethingCSomething{}$. In other words, now $f$ is a mapping to a 3-vector
counting how often each character occurs in a string. Since extracting values
from a vector is tiresome, we will use the shorthand notation
$\TransitionCount{a}$ to refer to the number of letters a that appear. E.g.
$\TransitionCount{a}(\text{curious and curiouser still}) = 1$. Note that we
still have the property that $\TransitionCount{o}(\text{homomorphism}) =
\TransitionCount{o}(\text{homo}) + \TransitionCount{o}(\text{morphism}) = 2 + 1
= 3$. If you treat the addition as element-wise vector addition, this approach
works for the vector version as well.

For our constraint, we will use $\TransitionCount{a} > \TransitionCount{b}$,
that is finding a count of characters in a string accepted by both automata where
there are more a:s than b:s. Since the underlying calculus is the same, we start immediately from \cref{fig:example-simplify-1}.

  We now need to relate the number of times a transition was taken to the
  character counts we are really interested in, as opposed to just the length in
  the previous example. For the $\Sigma$ labels, since we do not know which
  letter to use, we introduce a sum $r_{a1} + r_{b1} + r_{c1}$ per $\Sigma$
  transition.

  As in the previous example, we sum the occurrences for each transition to
  obtain the counts, here with the element-wise vector addition unpacked:
  \[
\MonoidElement =   \begin{bmatrix}
  r_{a1} + r_{a2} \\
  r_{b1} + r_{b2} \\
  r_{c1} + r_{c2} + 1 \\
  \end{bmatrix} =
  \begin{bmatrix}
    2 -2l_c' \\
    2l_b' + l_c' \\
    l_c + l_c' \\
    \end{bmatrix} =
    \begin{bmatrix}
      \TransitionCount{a} \\
      \TransitionCount{b} \\
      \TransitionCount{c}
      \end{bmatrix}
  \]

These equalities can be combined with the constraint that $\TransitionCount{a} >
\TransitionCount{b}$ to obtain $2 - l_c' > l_b'$. Since we have $1-l_c' \geq 0$
from one of the transition labels, we know $l_c' \geq 0$, and therefore that $1
> l_b' \implies l_b' = 0$. Plugging in the same inequality again on the same
principle, we get $2 - 2l_c' > l_c' \iff l_c' = 0$. This gives us the much
smaller automaton of \cref{fig:example-simplify-3}.

% \begin{figure}[t]
%   \centering 
%     \includegraphics[width=0.70\linewidth]{aca_or_bc_simplified_2}
%     \caption{$\AcaOrBc{}$ after removing the now unused b-transition.}\label{fig:example-simplify-2}
%     \Description[$\AcaOrBc$ without the returning transition from $B$ to $S$]{$\AcaOrBc$ without the returning transition from $B$ to $S$, which now makes the transition from $S$ to $B$ also have the same associated variable $l_c'$.}
%   \end{figure}

This will produce a straightforward product since both automata now have the
same shape, where we essentially only need to expand the self-loops of
$\SomethingCSomething$ and reduce its range labels to only capture a's, as seen
in \cref{fig:final-example}. Note that we got this result by only performing
linear inequality reasoning on the automata of the product versus the number of
times each transition would be used.

%   \begin{figure}[t]
% \centering
%     \includegraphics[width=0.75\linewidth]{original_product}
%     \caption{$\AcaOrBc{} \times \SomethingCSomething{}$ as it would have
%     appeared if we had computed it from the initial
%     automata.}\label{fig:original-product}
%   \end{figure}

To obtain the final Parikh image, we again need to tie the inequalities for the
transitions of the automata of the product to the new transition counters we
introduced in the product automaton. The only interesting transition is the one
carrying the c, so we will choose that one for illustration. In that case, the
bridging equation is $r_2 = l_c \land r_2 = r_{\Sigma}$ (and for all others
$=1$), which gives the Parikh image $\exists_{r_2} \TransitionCount{a} = 1 \land
\TransitionCount{b} = 0 \land \TransitionCount{c} = r_2 + 1$, and after
existence-elimination using reasoning on lower bounds (e.g. $r_2 \geq 0$) we
arrive at $\TransitionCount{c} \geq 1$.

>>>>

\usepackage{amsmath,empheq,fancybox}
\usepackage{paralist}
\usepackage{url}
\usepackage{color}
\usepackage{textcomp,listings}
\usepackage{array}
\usepackage{mymacros}
\usepackage{microtype}
\usepackage{listings}
\usepackage{csquotes}
\usepackage{proof}
\usepackage[capitalise]{cleveref}
\usepackage{algorithm2e}      
\usepackage{multirow}
\usepackage{mathpartir}
\usepackage{amsthm}
\usepackage{numprint}
\usepackage{ebproof}
\usepackage[makeroom]{cancel}

@@
